\textsc{Solution 1:}
Let $n \geq 1$.
We want to show 
\[
B_n = \{a^k \,|\, k \text{ is a multiple of $n$} \}
\] 
is regular. 
Define the following DFA $M$ (over $\Sigma = \{a\}$) as follows:
\begin{center}
\begin{tikzpicture}[shorten >=1pt,node distance=2cm,auto,initial text=]
\node[state,initial,accepting] (A) at (  0,  0) {$q_0$};
\node[state] (B) at (2, 0) {$q_1$};
\node[state] (C) at (4, 0) {$q_2$};
\draw(6, 0) node (D) {$...$};
\node[state] (E) at (8, 0) {$q_{n-1}$};

\path[->]
(A) edge [] node {$a$} (B)
(B) edge [] node {$a$} (C)
(C) edge [] node {$a$} (D)
(D) edge [] node {$a$} (E)
(E) edge [bend right=40,above] node {$a$} (A)
;
\end{tikzpicture}
\end{center}
Formally the DFA $M$ is defined as
$M = (\Sigma, Q, q_0, \delta, F)$ where:
\begin{enumerate}[topsep=0in,parsep=0in]
\item $\Sigma = \{a\}$
\item $Q = \{q_0, q_1, \ldots, q_{n-1}\}$
\item $\delta : Q \times \Sigma \rightarrow Q$ is defined by 
 \[
 \delta(q_i, a)
 =
 \begin{cases}
 q_{i + 1} & \text{if } 0 \leq i < n - 1 \\
 q_0       & \text{if } i = n - 1
 \end{cases}
 \]
\item $F = \{q_0\}$
\end{enumerate}
Clearly the strings accepted by $M$ are $\epsilon$, $a^n$, $a^{2n}$, $\ldots$,
i.e. $L(M) = \{a^k \,|\, \text{$k$ is a multiple of $n$}\} = B_n$.
Hence $B_n$ must be regular.

\textsc{Solution 2:}
Let $n \geq 1$.
We want to show 
\[
B_n = \{a^k \,|\, k \text{ is a multiple of $n$} \}
\] 
is regular. Note that 

\begin{align*}
B_n 
&= \{a^k \,|\, k \text{ is a multiple of $n$} \} \\
&= \{a^{mn} \,|\, m \geq 0 \} \\
&= \{(a^n)^m \,|\, m \geq 0 \} \\
&= \{a^n\}^* \\
&= L(a^n)^* \\
&= L((a^n)^*) \\
\end{align*}
i.e. $B_n$ is the language accepted by the regular expression $r = (a^n)^*$.
We already know that the language generated by a regular expression is also
accepted by a DFA.
Hence $B_n$ must be regular.

{\textsc Note.}
Note that the second solution is a lot clearer since
$L(r) = \{a^k \mid k \text{ is a multiple of $n$} \}$ is
shown completely.
However in the first solution the statement
\[
L(M) = \{a^m \mid m \text{ is a multiple of $n$} \}
\]
is not so immediate and properly speaking requires some proof.
You can formally show that 
\[
L(M) = \{a^m \mid m \text{ is a multiple of $n$} \}
\]
using mathematical induction.
This is the reason why in CS and Math, we frequently have several different
ways of looking at the same concept.
(In the case of regular languages, a regular language is one that
is accepted by or DFA, \textit{or} is accepted by an NFA, \textit{or}
is generated by a regular expression.)
Sometimes one way of looking at a problem will yield a more natural 
solution/proof or one that is shorter.
