\input{thispreamble.tex}

\renewcommand\AUTHOR{John Doe}

\begin{document}
\topmatter

\textsc{Objectives}
\begin{itemize}
\li Design context-free grammars.
\li Design PDA.
\end{itemize}


For questions asking you show an operation on context-free languages
language is closed, you need not prove your construction is correct.
(But you are welcome to do so. Most of such proofs, if not immediate,
is by induction.)
If the question does not say so, you can choose to give a PDA or a
context-free grammar.
For a PDA, draw the PDA state diagram and give the formal definition.

Study the solutions to the following questions in the textbook:
2.3, 2.4 (a) and (d), 2.6(a) and (c), 2.7, 2.8.

Here are the questions you should work.
Some questions have solutions (either from
the book or I have written up the solution).
\begin{myenum}
\li Sipser 2.1. Q1. Solution to 2.1(b) is provided. Study it carefully. 
\li Sipser 2.2. Q2. You may assume that $\{a^nb^nc^n \mid n \geq 0\}$ is not a
context-free language.
\li Sipser 2.3. Solution is provided in the textbook. Study it carefully.
\li Sipser 2.4. Q3. Solutions to 2.4(a) and 2.4(d) are provided in the textbook.
Study it carefully.
\li Sipser 2.5. Q4.
\li Sipser 2.6. Q5. Solutions to 2.6(a) and 2.6(c) are provided in the textbook. Study it carefully.
\li Sipser 2.7. Solution to 2.7 is provided in the textbook. Study it carefully.
\li Sipser 2.8. Solution to 2.8 is provided in the textbook. Study it carefully.
\li Sipser 2.9. Q6.
\li Sipser 2.10. Q7.
\li Sipser 2.13. Q8. Solution to 2.13(a) is provided. 
\end{myenum}

%------------------------------------------------------------------------------
\newpage
\textsc{How to draw a state diagram}

Here's an example showing you how to draw the elements of a state diagram.
Also, look at the solution to 1.3 below.

\begin{center}
  \begin{tikzpicture}[shorten >=1pt,>=triangle 60,double distance=2pt,node distance=2cm,auto,initial text=]
    
    \node[state,initial]   (q0) at (0, 0) {$q_0$};
    \node[state]           (q1) at (4, 0) {$q_1$};
    \node[state,accepting] (q2) at (8, 0) {$q_2$};
    \node[state]           (q3) at (0,-4) {$q_3$};
    \node[state]           (q4) at (4,-4) {$q_4$};

    \node[state]           (q5) at (12, 0) {$q_5$};
    \node[state]           (q6) at (12,-4) {$q_6$};

    \path[->]
    (q0) edge                node {$1$} (q1)    
    (q1) edge [bend left=15] node {$0$} (q2)
    (q0) edge [left]  node {$\alpha$} (q3)
    (q1) edge [right] node {$\beta$}  (q4)
    (q1) edge [loop above]   node {$1$} (q1)
    (q2) edge [loop below]   node {$0$} ()
    (q2) edge [bend left=15] node {$1$} (q1)
    (q3) edge [loop right]   node {$0,1$} ()
    (q4) edge [loop left]    node {$2,3$} ()
    (q5) edge [right, bend left=30]    node {$\gamma$} (q6)
    (q5) edge [right, bend left=60]    node {$\ep$}    (q6)
    (q5) edge [left, bend right=30]    node {$\delta$} (q6)
    ;
  \end{tikzpicture}
\end{center}

For more information on drawing state diagrams go to my tutorials at
\url{http://yliow.github.io} and 
look for \verb!latex-automata.pdf!:
Let me know if you have any questions about drawing state diagram.


%------------------------------------------------------------------------------
\newpage
\textsc{How to write a context-free grammar}

Here's a context-free grammar, $G$, for our $\{ a^n b^n \mid n \geq 0 \}$:
\[
G:
\begin{cases}
S \rightarrow \ep \\
S \rightarrow aSb
\end{cases}
\]
Usually we combine the productions for the same variable like this:
\[
G:
S \rightarrow \ep \mid aSb
\]
Here's an example of a longer grammar:
\[
G_1:
\begin{cases}
S \rightarrow \ep \mid TU \\
T \rightarrow a \mid   aT \\
U \rightarrow \ep \mid bUcc
\end{cases}
\]
And remember to list the starting variable first.

%------------------------------------------------------------------------------
\newpage
Q1. Sipser 2.1. 

\textsc{Solution}.

\input{q01.tex}

%------------------------------------------------------------------------------
\newpage
Q2. Sipser 2.2.

\textsc{Solution}.

\input{q02.tex}

%------------------------------------------------------------------------------
\newpage
Q3. Sipser 2.4.(b), (c), (e), (f)

\textsc{Solution}.

\input{q03.tex}

%------------------------------------------------------------------------------
\newpage
Q4. Sipser 2.5. For each questions, draw PDA state diagram and
informally describe how it words.
You do \textit{not} need to give the formal definition.

\textsc{Solution}.

\input{q04.tex}

%------------------------------------------------------------------------------
\newpage
Q5. Sipser 2.6 (b) and (d).

\textsc{Solution}.

\input{q04.tex}

%------------------------------------------------------------------------------
\newpage
Q6. Sipser 2.9.

\textsc{Solution}.

\input{q05.tex}

%------------------------------------------------------------------------------
\newpage
Q7. Sipser 2.10. (Hint: Look at the word \lq\lq or".)

\textsc{Solution}.

\input{q07.tex}

%------------------------------------------------------------------------------
\newpage
Q8. Sipser 2.13.

\textsc{Solution}.

\input{q08.tex}

\end{document}
