\input{myassignmentpreamble}
\input{ciss362}
\input{yliow}

\renewcommand\AUTHOR{John Doe}
\renewcommand\EMAIL{jdoe@cougars.ccis.edu}
\renewcommand\TITLE{Assignment 2}

\begin{document}
\topmatter

You should study Sipser chapter 0 and do all the problems in the Exercises
and Problems section.
Form study groups. Meet at CS Hangout. Compare solutions.
As always I very strongly encourage discussion.
But you must must write your own work and you must understand what you wrote
down.
If I call on you to explain your work and you can't, then that's considered
plaigiarism.

For Q1, enter your work in \verb!q01.tex!.
Several examples are included so that you and modify the given \LaTeX\ code.
You can also find some information on \LaTeX\ at my website
\url{http://yliow.github.io} (go to the Tutorial section and look for
the latex pdf.) Or just ask me.

\newpage
\input{induction.tex}

%------------------------------------------------------------------------------
\newpage
Q1. Sipser 0.10. 

\SOLUTION

\input{q01.tex}

%------------------------------------------------------------------------------
\newpage
Q2. Sipser 0.12.

\SOLUTION

\input{q02.tex}

%------------------------------------------------------------------------------
\newpage
Q3.
Using mathematical induction, give a complete proof of the following fact:
\[
  4(1^3 + 2^3 + \cdots + n^3) = n^2(n+1)^2
\]
for $n \geq 0$.

\SOLUTION

\input{q03.tex}

%------------------------------------------------------------------------------
\newpage
Q4.
Using WOP, give a complete proof of the following fact:
\[
  4(1^3 + 2^3 + \cdots + n^3) = n^2(n+1)^2
\]
for $n \geq 0$.


\SOLUTION

\input{q04.tex}

%------------------------------------------------------------------------------
\newpage
Q5. Write a complete proof of the Euclidean property of integers:
If $a$ and $b$ are integers such that $a \geq 0$ and $b > 0$,
then there exists integers $q$ and $r$ such that $q \geq 0$ and
\[
a = bq + r, \,\,\, 0 \leq r < b
\]
You must use WOP.
(I have given you the proof in class.
The point of this question is to study the proof in depth, write it up
properly in your own words, and edit your work until it's perfect.) 

\SOLUTION

\input{q05.tex}

%------------------------------------------------------------------------------
\newpage
\input{primes}

%------------------------------------------------------------------------------
\newpage
Q\textred{6}.
Consider the following fact:
Every positive integer $n \geq 1$ is a product of primes.
For instance for $n = 20$,
the ordered collection of primes involved are $(2, 2, 5)$ (in ascending order).
We define the product of the empty collection of primes, i.e. $()$, to be $1$.
Of course the collection can have one single number:
If the collection is $(11)$, then the product is 11.

Define $P(n)$ to be the above statement, i.e.,
\[
  P(n) =
  \biggl(
  \text{Every positive integer $n \geq 1$ is a product of primes}
  \biggr)
\]
Give a complete proof of $P(n)$ for all $n \geq 1$
using strong or weak induction.
Here, I'm giving you the $P(n)$ for free.

You can assume Euclid's lemma: If $p$ is a prime dividing $ab$,
then $p$ divides $a$ or $p$ divides $b$.

(In fact not only is every positive integer $n$ representable
as a product of primes,
it is represented as a product of primes in a \textit{unique way}.
In other words, the ordered collection of primes for $n$ is unique.
This can also be proven using induction.
But you don't have to prove it.)

\SOLUTION

\input{q06.tex}

%------------------------------------------------------------------------------
\newpage
Q\textred{7}.
Using mathematical induction, 
prove that for $n \geq 4$, $2^n \geq n^2$.

(Actually this can also be proven using Calculus.)

\SOLUTION

\input{q07.tex}

%------------------------------------------------------------------------------
\newpage
Q\textred{8}. Prove or disprove the following: Let $X$ and $Y$ be sets.
\begin{enumerate}
  \item[(a)] $P(X \times Y) = P(X) \times P(Y)$
  \item[(b)] $P(X \cup Y) = P(X) \cup P(Y)$
  \item[(c)] $P(X) = 2^{|X|}$ if $X$ is a finite set.
\end{enumerate}
Remember that to disprove a statement, all you need is a
counterexample.
When constructing examples, always construct the simplest.

\SOLUTION

\input{q08.tex}

%------------------------------------------------------------------------------
\newpage
Q\textred{9}.
Using mathematical induction, prove that an undirected simple graph
with $n$ nodes can have at most $n(n + 1)/2$ edges.
(Simple means there are no loop edges and no multiedges.
A loop edge is an edge that joins the same node.
Two edges are multiedges is the join the same two nodes.)
$P(n)$ is the statement
\lq\lq if $G$ is an undirected simple graph with $n$ nodes, then the
$G$ has at most $n(n+1)/2$ edges".

\SOLUTION

\input{q09.tex}

%------------------------------------------------------------------------------
\newpage
Q\textred{10}.
A triangulation graph is a graph which looks like a patchwork of 
 triangles, i.e., 
 the graph is constructed by \lq\lq glueing'' triangles
 on a tabletop. 
\begin{center}
\begin{tikzpicture}[shorten >=1pt,node distance=2cm,auto,initial text=]
\node   (a) [shape=circle, draw] at (0, 0) {};
\node   (b) [shape=circle, draw] at (1, 2) {};
\node   (c) [shape=circle, draw] at (2, 0) {};
\path[-] (a) edge node {} (b)
(b) edge node {} (c)
(c) edge node {} (a)
;
\end{tikzpicture}
\end{center}
along edges.
 Note that line (edge) crossings are not allowed.
 For instance here's one 5 nodes (i.e. circles) and 7 edges (i.e. lines):

\begin{center}
\begin{tikzpicture}[shorten >=1pt,node distance=2cm,auto,initial text=]
\node   (a) [shape=circle, draw] at (0, 3) {};
\node   (b) [shape=circle, draw] at (2, 3) {};
\node   (c) [shape=circle, draw] at (4, 2.5) {};
\node   (d) [shape=circle, draw] at (1, 0) {};
\node   (e) [shape=circle, draw] at (3, 1) {};
\path[-] (a) edge node {} (b)
(b) edge node {} (c)
(a) edge node {} (d)
(b) edge node {} (d)
(b) edge node {} (e)
(d) edge node {} (e)
(c) edge node {} (e)
;
\end{tikzpicture}
\end{center}

(a) Is it true that using the above construction (i.e., glueing
triangle graphs along edges), all the \lq\lq pieces'' (i.e. finite regions) 
that you see 
are always triangles (i.e. a shape that is bounded by 
exactly 3 edges)? For instance in the above graph, there are 3 triangles.

(b) The degree of a node is just the number of edges (lines) joined to it.
For the above type of graphs is it possible to draw one 
with exactly one node of degree 10? 
Show that it is possible by drawing such a graph or prove it's not possible.

(c) Is it possible to have a graph with 5 nodes of degree 2? 
Show that it is possible by drawing such a graph or prove it's not possible.

(d) Is it possible to have a graph with no nodes of degree 2? 
Show that it is possible by drawing such a graph or prove it's not possible.

(e) The degrees of the nodes in the above example is 2, 2, 3, 3, 4.
Is it possible to have a graph with degrees 2, 2, 2, 2, 4?
Show that it is possible by drawing such a graph or prove it's not possible.

(f) The degrees of the nodes in the above example is 2, 2, 3, 3, 4.
Is it possible to have a graph with degrees 2, 2, 3, 3, 3?
Show that it is possible by drawing such a graph or prove it's not possible.

(g) Is it possible to have a triangulation with no nodes of degree 2 or 3?

(h) Suppose you have a triangular graph $G$ with a node $v$ of degree 2.
If there are at least 4 nodes in $G$, after removing $v$ from $G$
and removing all the edges attached to $v$, is the resulting graph
a triangulation graph? 
Prove that this is true or provide a counterexample.
       
\SOLUTION

\input{q10.tex}

%------------------------------------------------------------------------------
\newpage
Q\textred{11}.
Sipser 0.13.

\SOLUTION

\input{q11.tex}

\end{document}
