\input{thispreamble.tex}

\renewcommand\AUTHOR{John Doe}

\begin{document}
\topmatter

\textsc{Objectives}
\begin{itemize}
\li Design DFAs.
\li Design NFAs.
\li Design regexes.
\li Show a language is regular by designing a DFA, NFA, or regex.
\li Show a language construction (operator) is a closed operator on
regular languages.
\end{itemize}

For questions asking you show an operation on regular
language is closed, you need not prove your construction is correct.
(But you are welcome to do so. Most of such proofs, if not immediate,
is by induction.)
Drawing a diagrma usually helps.
For solutions provided, make sure you study it carefully
as a guide for answering such questions.


\begin{myenum}
\li Sipser 1.31. Solution provided. Study it carefully. This is from a06. 
\li Sipser 1.32. Q1
\li Sipser 1.33. Q2
\li Sipser 1.34. Q3
\li Sipser 1.35. Q4
\li Sipser 1.36. Solution provided. Study it carefully.
\li Sipser 1.37. Q5
\li Sipser 1.38. Q6
\li Sipser 1.40. Q7. Solution to 1.40(a) is in the textbook. Study it carefully.
\li Sipser 1.41. Q8
\li Sipser 1.42. Q9
\end{myenum}

%------------------------------------------------------------------------------
\newpage
\textsc{How to draw a state diagram}

Here's an example showing you how to draw the elements of a state diagram.
Also, look at the solution to 1.3 below.

\begin{center}
  \begin{tikzpicture}[shorten >=1pt,>=triangle 60,double distance=2pt,node distance=2cm,auto,initial text=]
    
    \node[state,initial]   (q0) at (0, 0) {$q_0$};
    \node[state]           (q1) at (4, 0) {$q_1$};
    \node[state,accepting] (q2) at (8, 0) {$q_2$};
    \node[state]           (q3) at (0,-4) {$q_3$};
    \node[state]           (q4) at (4,-4) {$q_4$};

    \node[state]           (q5) at (12, 0) {$q_5$};
    \node[state]           (q6) at (12,-4) {$q_6$};

    \path[->]
    (q0) edge                node {$1$} (q1)    
    (q1) edge [bend left=15] node {$0$} (q2)
    (q0) edge [left]  node {$\alpha$} (q3)
    (q1) edge [right] node {$\beta$}  (q4)
    (q1) edge [loop above]   node {$1$} (q1)
    (q2) edge [loop below]   node {$0$} ()
    (q2) edge [bend left=15] node {$1$} (q1)
    (q3) edge [loop right]   node {$0,1$} ()
    (q4) edge [loop left]    node {$2,3$} ()
    (q5) edge [right, bend left=30]    node {$\gamma$} (q6)
    (q5) edge [right, bend left=60]    node {$\ep$}    (q6)
    (q5) edge [left, bend right=30]    node {$\delta$} (q6)
    ;
  \end{tikzpicture}
\end{center}

For more information on drawing state diagrams go to my tutorials and
look for \verb!latex-automata.pdf!:
\begin{center}
{\small \url{https://drive.google.com/file/d/1AeE-P0WNvQlitzPDxQpGE8bMR9Yc9gMW}}
\end{center}
Let me know if you have any questions about drawing state diagram.

%------------------------------------------------------------------------------
\newpage
Sipser 1.31. (This is from a06. Solution provided.) 

\textsc{Solution}.

\input{sipser3-1-31.tex}

%------------------------------------------------------------------------------
\newpage
Q1. Sipser 1.32.

\textsc{Solution}.

\input{q01.tex}

%------------------------------------------------------------------------------
\newpage
Q2. Sipser 1.33.

\textsc{Solution}.

\input{q02.tex}

%------------------------------------------------------------------------------
\newpage
Q3. Sipser 1.34.

\textsc{Solution}.

\input{q03.tex}

%------------------------------------------------------------------------------
\newpage
Q4. Sipser 1.35.

\textsc{Solution}.

\input{q04.tex}

%------------------------------------------------------------------------------
\newpage
Sipser 1.36. Solution provided. Study it carefully.

(Hint: In this case, it's easier not to use automata construction.
Use a regular expression instead.)

\textsc{Solution}.
 
\input{sipser3-1-36.tex}

%------------------------------------------------------------------------------
\newpage
Q5. Sipser 1.37.

\textsc{Solution}.

\input{q05.tex}

%------------------------------------------------------------------------------
\newpage
Q6. Sipser 1.38.

\textsc{Solution}.

\input{q06.tex}

%------------------------------------------------------------------------------
\newpage
Q7. Sipser 1.40. Solution to 1.40(a) is in the textbook -- study it carefully.

\textsc{Solution}.

\input{q07.tex}

%------------------------------------------------------------------------------
\newpage
Q8. Sipser 1.41.

\textsc{Solution}.

\input{q08.tex}

%------------------------------------------------------------------------------
\newpage
Q9. Sipser 1.42.

\textsc{Solution}.

\input{q09.tex}

\end{document}
